---
title: "Machine Learning Chapter 9 Homework"
author: "Yifan YE"
date: "11/24/2018"
output: html_document
---
## Chapter 9 Problem 3

In this question, we explore the maximal margin classifier.

(a). We have 7 observations from the descriptions of the question. They have been assigned to two groups (according to their colors) and we plot them below (using symbols from textbook):

```{r scatter plot}
# scatter plot
x1 = c(3, 2, 4, 1, 2, 4, 4)
x2 = c(4, 2, 4, 4, 1, 3, 1)
col = c("red", "red", "red", "red", "blue", "blue", "blue")
plot(x1, x2, col = col, xlim = c(-1, 6), ylim = c(0, 4))
```

(b). Here we find the optimal separating hyperplane of the form $\beta_0+\beta_1X_1+\beta_2X_2=0$. Notice that the points that are most close to the desired hyperplane are $(2,2), (4,4), (2,1), (4,3)$. The slope between those two red points is the same as the slope between those two blue points (both are 1). So this leads to that the slope of the optimal separating hyperplane (here it is a straight line) is also 2 because the optimal separating hyperplane must keep a banlance with both groups of data. Here the line corresponding to red points is $X_2=X_1$ or $X_1-X_2=0$. The line corresponding to blue points is $X_2=X_1-1$ or $X_1-X_2-1=0$. So we take the "meddle" line between them and get the form of the desired htperplane is $X_2=X_1-0.5$ or $X_1-X_2-0.5=0$. Otherwise the min distance between sample points to the plane is less than the case before due to the straight line rotate a little bit and make the min distance shorter. 

Just to mention a little bit, the problem can be written into:
$$\max_{\beta_0,\beta_1,\beta_2}M$$ 
under the constraints:
$$\beta_0^2+\beta_1^2+\beta_2^2=1$$
and:
$$y_i(\beta_0+\beta_1x_{i1}+\beta_2x_{i2})\geq M,\ \forall i$$

We can also solve the programming problem we get the same result as before.

```{r optimal separating hyperplane}
# optimal separating hyperplane 
# we get the line's form from above
plot(x1, x2, col = col, xlim = c(-1, 6), ylim = c(0, 4))
abline(-0.5, 1) # slope = 1, intercept = -0.5
```

(c). From basic math we know that the "red" group in the plot corresponds to the "upper" side of the straight line, which implies $X_2>X_1-0.5$ or $X_1-X_2-0.5<0$. So classifying rule is that new point should be classified to "red" if $X_1-X_2-0.5<0$ or $0.5-X_1+X_2>0$.

(d). As we have said before, margins are determined by support vectors and hence their forms are $X_1-X_2-1=0$ and $X_1-X_2=0$. Margins are plotted by dashed lines below.

```{r margins}
# margins
plot(x1, x2, col = col, xlim = c(-1, 6), ylim = c(0, 4))
abline(-0.5, 1)
abline(-1, 1, lty = 4)
abline(0, 1, lty = 4)
```

(e). We first find those closest points' projections on the line. 
$$(2,2)->(2.25,1.75), (4,4)->(4.25,3.75), (2,1)->(1.75,1.25), (4,3)->(3.75,3.25)$$ by calculating the intersection point of the line and a line that is perpendicular to it and also pass through the point. Plots are displayed below:

```{r support vectors}
# support vectors
plot(x1, x2, col = col, xlim = c(-1, 6), ylim = c(0, 4))
abline(-0.5, 1)
abline(-1, 1, lty = 4)
abline(0, 1, lty = 4)
arrows(2, 1, 1.75, 1.25)
arrows(2, 2, 2.25, 1.75)
arrows(4, 4, 4.25, 3.75)
arrows(4, 3, 3.75, 3.25)
```

(f). First we need to clarify that the $7^{th}$ point $(4,1)$ is not a sopport vector (out of the range of margins). Because the maximal margin hyperplane depends directly on the support vectors, but not on the other observations or we can see from above directely that if we change the location of the right bottom point then the result will not change. So we can say that a slight movement of the seventh observation would not affect the maximal margin hyperplane.

(g). We can just slight rotate the optimal one $X_2=X_1-0.5$ a little bit and get $X_2=0.8X_1$. Plot is below.

```{r not optimal} 
# not optimal
plot(x1, x2, col = col, xlim = c(-1, 6), ylim = c(0, 4))
abline(-0.5, 1)
abline(0, 0.8, col="red")
```

(h). We just need to add a blue points to the inside of the red group.

```{r new points}
# new points
x3 = c(3, 2, 4, 1, 2, 4, 4, 2)
x4 = c(4, 2, 4, 4, 1, 3, 1, 4)
col2 = c("red", "red", "red", "red", "blue", "blue", "blue", "blue", "blue", "red", "red", "red", "blue")
plot(x3, x4, col = col2, xlim = c(-1, 6), ylim = c(0, 4))
```

